### 2015-08-10 conclusion.tex
Minor tweaks, Final version.

The complexity of managing HPC software is rapidly increasing, and it
will continue unabated without better tools.
In this paper, we reviewed the state of software management tools
across a number of HPC sites, with particular focus on Livermore
Computing (LC). While tools exist that can handle multi-configuration
installs, none of them addresses the combinatorial nature
of the software configuration space directly. None of them allows
a user to {\it compose} new builds with version, compiler,
and dependency parameters rapidly.

>>>>

The complexity of managing HPC software is rapidly increasing, and it
will continue unabated without better tools.
In this paper, we reviewed the state of software management tools
across a number of HPC sites, with particular focus on Livermore
Computing (LC). While existing tools can handle multi-configuration
installs, none of them sufficiently addresses the combinatorial nature
of the software configuration space. None of them allows
a user to {\it compose} new builds with version, compiler,
and dependency parameters rapidly.

### 2015-08-10 implementation-environment.tex
Minor tweaks, Final version.

\subsubsection{Environment isolation}
Inconsistencies arise not only from building manually but also due to 
package installation and user environment differences.
%
For example, LLNL performance tools use two versions of the {\tt libelf}
library. One is distributed with RedHat Linux, while the
other is publicly available. They have the same API but an incompatible ABI.
Specifying the wrong version at build time has caused many
unexpected crashes.

@@

\subsubsection{Compiler wrappers and RPATHs}
Finding dependencies at build time is not the only obstacle to reproducible
behavior.  As mentioned in Section~\ref{sec:motivation}, binaries must be 
able to find dependency libraries at {\it runtime}.
One of the most common user errors at LLNL is improper library configuration.
Users frequently do not know the libraries with which a package was built, and
constructing a suitable {\tt LD\_LIBRARY\_PATH} for a package that was built 
by someone else is difficult. To reduce the number of support calls that we 
receive, we typically add {\tt RPATHs} to public software installations, so 
that paths to dependencies are embedded in binaries. Thus, users do not need 
to know details of dependencies to run installed software correctly.

>>>>

\subsubsection{Environment Isolation}
Inconsistencies arise not only from building manually but also due to 
differences between the package installation environment and the user
environment.
%
For example, LLNL performance tools use two versions of the {\tt libelf}
library. One is distributed with RedHat Linux, while the
other is publicly available. They have the same API but an incompatible ABI.
Specifying the wrong version at build time has caused many
unexpected crashes at runtime.

@@

\subsubsection{Compiler Wrappers and RPATHs}
Finding dependencies at build time is not the only obstacle to reproducible
behavior.  As mentioned in Section~\ref{sec:motivation}, binaries must be 
able to find dependency libraries at {\it runtime}.
One of the most common user errors at LLNL is improper library configuration.
Users frequently do not know the libraries with which a package was built, and
constructing a suitable {\tt LD\_LIBRARY\_PATH} for a package that was built 
by someone else is difficult. To reduce the number of support calls that we 
receive, we typically add {\tt RPATHs} to public software installations, so 
that paths to dependencies are embedded in binaries. Thus, users do not need 
to know details of dependencies to run installed software correctly.

### 2015-08-10 implementation-specs.tex
Minor tweaks, Final version.

\subsubsection{Configuration Complexity}
A spec DAG has many degrees of freedom, and users cannot reasonably be
expected to understand or to specify all of them.  In our experience at LLNL,
the typical user only cares about a small number of build constraints (if any),
and does not know enough to
specify the rest. For example, a user may know that a certain version of a library like
{\tt boost}~\cite{boost} is required, but only cares that other build parameters are set so that
the build will succeed.
%
Configuration complexity makes the HPC software ecosystem difficult to manage:
too many parameters exist to know all of them. However, the known and
important ones often provide detailed build constraints. Thus, we have two
competing concerns.  We need the ability to specify details
without having to remember all of them.

@@

We have developed a syntax for specs that allows users to specify
only the constraints that matter. Our syntax can represent software
DAGs concisely enough to use it on the command
line. The spec syntax is recursively defined to allow users to specify
parameters on dependencies as well as on the root of the DAG.
Figure~\ref{fig:grammar} shows the EBNF grammar through which Spack
flexibly supports constraints.

@@

\subsubsection{Constraints in packages}\label{sec:constraints}

@@

\subsubsection{Build specialization}

>>>>

\subsubsection{Configuration Complexity}
A spec DAG has many degrees of freedom, and users cannot reasonably be
expected to understand or to specify all of them.  In our experience at LLNL,
the typical user only cares about a small number of build constraints (if any),
and does not know enough to
specify the rest. For example, a user may know that a certain version of a library like
{\tt boost}~\cite{boost} is required, but only cares that other build parameters are set so that
the build will succeed.
%
Configuration complexity makes the HPC software ecosystem difficult to manage:
too many parameters exist to specify them all. However, the known and
important ones often provide detailed build constraints. Thus, we have two
competing concerns.  We need the ability to specify details
without having to remember all of them.

@@

We have developed a syntax for specs that allows users to specify
only constraints that matter to them. Our syntax can represent
DAGs concisely enough for command line use.
The spec syntax is recursively defined to allow users to specify
parameters on dependencies as well as on the root of the DAG.
Figure~\ref{fig:grammar} shows the EBNF grammar through which Spack
flexibly supports constraints.

@@

\subsubsection{Constraints in Packages}\label{sec:constraints}

@@

\subsubsection{Build Specialization}

### 2015-08-10 introduction.tex
Minor tweaks, Final version.

To support the diverse needs of applications, system administrators
and developers frequently build, install, and support many
different configurations of math and physics libraries, and
other software.  Frequently, applications must be rebuilt to fix bugs
and to support new versions of the operating system (OS), 
MPI implementation, compiler, and other
dependencies.  Unfortunately, building scientific software is
notoriously complex, with immature build systems that are difficult to
adapt to new
machines~\cite{dubois+:comp-sci-eng,hoste+:pyhpc12,wilson+:corr}.
Worse, the space of required builds grows combinatorially with each
new configuration parameter. As a result, LLNL staff spend countless
hours dealing with build and deployment issues.

>>>>

To support the diverse needs of applications, system administrators
and developers frequently build, install, and support many
different configurations of math and physics libraries, as well as
other software.  Frequently, applications must be rebuilt to fix bugs
and to support new versions of the operating system (OS), 
MPI implementation, compiler, and other
dependencies.  Unfortunately, building scientific software is
notoriously complex, with immature build systems that are difficult to
adapt to new
machines~\cite{dubois+:comp-sci-eng,hoste+:pyhpc12,wilson+:corr}.
Worse, the space of required builds grows combinatorially with each
new configuration parameter. As a result, LLNL staff spend countless
hours dealing with build and deployment issues.

### 2015-08-10 limitations.tex
Minor tweaks, Final version.

Spack's current implementation has several limitations.
As mentioned in Section~\ref{sec:concretization}, we use a greedy 
concretization algorithm. For example, if package {\tt P} depends on 
both {\tt hwloc@1.9} and {\tt mpi}, and if the algorithm
chooses an MPI implementation for {\tt P} that depends strictly on
{\tt hwloc@1.8}, a conflict on {\tt hwloc} arises. In this case, Spack 
raises an error and the user must resolve the issue. Our implementation 
does not backtrack to find an MPI version that does not conflict. These 
cases have been rare so far. However, we plan to add better constraint
satisfaction solving to Spack in future work.

@@

In the near term, we are actively working to grow a community around Spack 
and to build a larger base of contributors. We plan to
increase build robustness by deploying a continuous testing system
at LLNL.  Finally, to support the growing number of
HPC languages, runtimes, and programming models, we will add
capabilities to Spack that allow packages to depend on particular
compiler features.  More and more, our codes are relying
on advanced compiler capabilities, like {\tt C++11} language features,
OpenMP versions, and GPU compute capabilities. Ideally, Spack
will find suitable compilers and ensure ABI consistency when many
such features are in use.

>>>>

Spack's current implementation has several limitations.
As mentioned in Section~\ref{sec:concretization}, we use a greedy 
concretization algorithm. For example, if package {\tt P} depends on 
both {\tt hwloc@1.9} and {\tt mpi}, and if the algorithm
chooses an MPI implementation for {\tt P} that depends strictly on
{\tt hwloc@1.8}, a conflict on {\tt hwloc} arises. In this case, Spack 
raises an error and the user must resolve the issue. Our implementation 
does not backtrack to find an MPI version that does not conflict. These 
cases have been rare so far. However, we plan to add better constraint
solving to Spack in future work.

@@

In the near term, we are actively working to grow a community around Spack 
and to build a larger base of contributors. We plan to
increase build robustness by deploying a continuous testing system
at LLNL.  Finally, to support the growing number of
HPC languages, runtimes, and programming models, we will add
capabilities to Spack that allow packages to depend on particular
compiler features.  More and more, our codes are relying
on advanced compiler capabilities, like C++11 language features,
OpenMP versions, and GPU compute capabilities. Ideally, Spack
will find suitable compilers and ensure ABI consistency when many
such features are in use.

### 2015-08-10 motivation.tex
Minor tweaks, Final version.

\paragraph{Traditional package managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
Anaconda~\cite{anaconda,conda}, uses the same approach but
is designed to run on top of a host OS.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package has only a single version,
and most install packages in a single, inflexible location.
To install multiple configurations, users must create custom, combinatorial
naming schemes to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

@@

Packaging problems arise in HPC because a supercomputer's hardware, OS, and
file system are shared by many users with different requirements.  The classic
solution to this problem is to use virtual
machines (VMs)~\cite{barham2003xen,rosenblum1999vmware,smith2005architecture}
or lightweight virtualization techniques like Linux
containers~\cite{felter2014updated,merkel2014docker}. This model allows each
 user to have a personalized environment with its own package manager, and it
has been extremely successful for servers at cloud data centers. VMs typically
have near-native compute performance but low-level HPC network drivers still
exhibit major performance issues. VMs are not well supported on many
non-Linux operating systems, an issue for the lightweight
kernels of bleeding-edge Blue Gene/Q and Cray machines.
Finally, each VM still uses a traditional package manager,
so running many configurations still requires a large number of VMs.
For facilities, this profusion of VMs is a security concern since it
complicates mandatory patching. Also, managing  a large number of VM
environments is tedious for users.

\begin{table*}\centering
\begin{tabular}{|l|l|}
\hline
Site           & Naming Convention \\
\hline
\hline
LLNL       & {\tt / usr / global / tools / \$arch / \$package / \$version} \\
           & {\tt / usr / local~ / tools / \$package-\$compiler-\$build-\$version } \\
\hline
Oak Ridge~\cite{jones+:cug08}  & {\tt / \$arch / \$package / \$version / \$build} \\
\hline
TACC/Lmod~\cite{mclay:lmod-tutorial}& {\tt / \$compiler-\$comp\_version / \$mpi / \$mpi\_version / \$package / \$version} \\
\hline
\hline
Spack default                  & {\tt / \$arch / \$compiler-\$comp\_version / \$package-\$version-\$options-\$hash} \\
\hline
\end{tabular}
\caption{
	Software organization of various HPC sites.
	\label{tab:naming-conventions}
}
\end{table*}

@@

The EasyBuild~\cite{hoste+:pyhpc12} tool is in production use at
the University of Ghent and the J\"ulich Supercomputing Center.
It allows multiple versions to be installed at once.  Rather
than setting {\tt RPATHs}, it auto-generates module files
to manage the environment, and it is closely coupled with
Lmod~\cite{geimer+:hust14}.  EasyBuild groups the compiler, MPI, FFT, and
BLAS libraries together in a {\it toolchain} that can be used by
package files. The grouping provides some composability and
separates compiler flags and MPI concerns from client packages.

HashDist~\cite{hashdist} is a meta-build system and package manager
for HPC.  Of the existing solutions, it is the most similar to Spack.
Like Nix, it uses cryptographic versioning and stores installations in
unique directories.
%
Both Nix and HashDist use {\tt RPATHs} in their packages to ensure that
libraries are found correctly.

\paragraph{Gaps in current practice}
The flexible cryptographic versioning of Nix and HashDist manages
the package {\it and} its dependency configuration and can represent any
configuration. However, users cannot easily navigate or query the installed
software.
%The systems are in a sense ``write-only'':
%Nix does not offer a way to view a package's dependencies.
%
EasyBuild and Smithy generate environment modules, which supports some
querying. Naming schemes used in existing module systems, however, cannot
handle combinatorial versions, which the Lmod authors call the
``matrix problem''~\cite{mclay:lmod-tutorial}.

@@

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatorial, and arbitrary combinations
of compiler, MPI version, build options, and dependency versions require
non-trivial modifications to many package files. Indeed, the number of package
files required for most existing systems scales with the number of version
{\it combinations}, not the number of packages, which quickly becomes
unmanageable.  As an example, the EasyBuild system has over 3,300 files for
several permutations of only 600 packages. A slightly different dependency
graph requires an entire new package file hierarchy.  HashDist supports
composition more robustly but does not have first-class parameters for
versions, compilers, or versioned interfaces.
%
HPC sites need better ways to {\it parameterize} packages so that new
builds can be {\it composed} in response to user needs.
%

>>>>

\paragraph{Traditional Package Managers}
Package managers automate the installation of complex sets of software packages.
{\it Binary package managers} such as RPM, yum, and APT~\cite{foster+:rpm03,silva:apt01,yum} are integrated with most
OS distributions, and they are used to ensure that dependencies
are installed before packages that require them.
Anaconda~\cite{anaconda,conda}, uses the same approach but
is designed to run on top of a host OS.
These tools largely solve the problem of managing a {\it single} software
stack, which works well for the baseline OS and drivers, which are
common to all applications on a system.
These tools assume that each package has only a single version,
and most install packages in a single, inflexible location.
To install multiple configurations, users must create custom, combinatorial
naming schemes to avoid conflicts. They typically require root
privileges and do not optimize for specific hardware.

@@

Packaging problems arise in HPC because a supercomputer's hardware, OS, and
file system are shared by many users with different requirements.  The classic
solution to this problem is to use virtual
machines (VMs)~\cite{barham2003xen,rosenblum1999vmware,smith2005architecture}
or lightweight virtualization techniques like Linux
containers~\cite{felter2014updated,merkel2014docker}. This model allows each
 user to have a personalized environment with its own package manager, and it
has been very successful for servers at cloud data centers. VMs typically
have near-native compute performance but low-level HPC network drivers still
exhibit major performance issues. VMs are not well supported on many
non-Linux operating systems, an issue for the lightweight
kernels of bleeding-edge Blue Gene/Q and Cray machines.
Finally, each VM still uses a traditional package manager,
so running many configurations still requires a large number of VMs.
For facilities, this profusion of VMs is a security concern because it
complicates mandatory patching. Also, managing  a large number of VM
environments is tedious for users.

\begin{table*}\centering
\begin{tabular}{|l|l|}
\hline
Site           & Naming Convention \\
\hline
\hline
LLNL & {\tt / usr / global / tools / \$arch / \$package / \$version} \\
           & {\tt / usr / local~ / tools / \$package-\$compiler-\$build-\$version } \\
\hline
ORNL~\cite{jones+:cug08}  & {\tt / \$arch / \$package / \$version / \$build} \\
\hline
TACC / Lmod~\cite{mclay:lmod-tutorial}& {\tt / \$compiler-\$comp\_version / \$mpi / \$mpi\_version / \$package / \$version} \\
\hline
\hline
Spack default                  & {\tt / \$arch / \$compiler-\$comp\_version / \$package-\$version-\$options-\$hash} \\
\hline
\end{tabular}
\caption{
	Software organization of various HPC sites.
	\label{tab:naming-conventions}
}
\end{table*}

@@

The EasyBuild~\cite{hoste+:pyhpc12} tool is in production use at
the University of Ghent and the J\"ulich Supercomputing Center.
It allows multiple versions to be installed at once.  Rather
than setting {\tt RPATHs}, it generates module files
to manage the environment, and it is closely coupled with
Lmod~\cite{geimer+:hust14}.  EasyBuild groups the compiler, MPI, FFT, and
BLAS libraries together in a {\it toolchain} that can be used by
package files. The grouping provides some composability and
separates compiler flags and MPI concerns from client packages.

HashDist~\cite{hashdist} is a meta-build system and package manager
for HPC.  Of the existing solutions, it is the most similar to Spack.
Like Nix, it uses cryptographic versioning and stores installations in
unique directories.
%
Both Nix and HashDist use {\tt RPATHs} in their packages to ensure that
libraries are found correctly at runtime.

\paragraph{Gaps in Current Practice}
The flexible cryptographic versioning of Nix and HashDist manages
the package {\it and} its dependency configuration and can represent any
configuration. However, users cannot easily navigate or query the installed
software.
%The systems are in a sense ``write-only'':
%Nix does not offer a way to view a package's dependencies.
%
EasyBuild and Smithy generate environment modules, which supports some
querying. Naming schemes used in existing module systems, however, cannot
handle combinatorial versions, which the Lmod authors call the
``matrix problem''~\cite{mclay:lmod-tutorial}.

@@

The main limitation of existing tools is the lack of build {\it composability}.
The full set of package versions is combinatorial, and arbitrary combinations
of compiler, MPI version, build options, and dependency versions require
non-trivial modifications to many package files. Indeed, the number of package
files required for most existing systems scales with the number of version
{\it combinations}, not the number of packages, which quickly becomes
unmanageable.  As an example, the EasyBuild system has over 3,300 files for
several permutations of around 600 packages. A slightly different dependency
graph requires an entire new package file hierarchy.  HashDist supports
composition more robustly but does not have first-class parameters for
versions, compilers, or versioned interfaces.
%
HPC sites need better ways to {\it parameterize} packages so that new
builds can be {\it composed} in response to user needs.
%

### 2015-08-10 usecase-ares.tex
Minor tweaks, Final version.

We have configured Spack to build ARES with external MPI implementations, 
depending on the host system. This configuration exploits the vendor- 
or site-supplied MPI installation that often uses host-specific optimized 
network drivers. MPI is shown as a virtual dependency in the figure, as 
the implementation differs according to the host machine.  ARES builds its 
own Python version in order to run on machines where Python is not well 
supported, like Blue Gene/Q.  In particular, ARES builds a version of 
Python 2.7 for Blue Gene/Q, which the native software stack does not support.

Prior to using Spack, ARES managed its software stack with MixDown.
Thus, the ARES team already had some experience supporting 
automated builds of dependencies. We developed Spack packages for the LLNL 
packages in Figure~\ref{fig:ares}. Many of the external packages were already 
available in Spack, but some, such as Python, required modifications to support
the new platforms and compilers.

@@

Altogether, the initial packaging effort required roughly two months 
for an experienced build engineer working 20 hours per week.
As shown in the table, 36 different configurations have been run using Spack
(up to 4 versions on each of 10 architecture-compiler-MPI combinations).
Prior to using Spack, only Linux/Intel configurations were automated. The ARES 
team listed a number of key features that enabled the increased automation:
\begin{enumerate}
\item Spack's version tracking and optional dependencies were required to
      build the four configurations with correct libraries;
\item The spec syntax allowed build scripts to test compiler,
      compiler version, and dependency versions concisely---a necessity
      for handling the different architectures;
\item Patching packages for particular platforms was
      necessary to build many packages; and
\item Using a DSL embedded in Python was a significant benefit;
      certain packages required custom scripting to patch.
\end{enumerate}

>>>>

We have configured Spack to build ARES with external MPI implementations, 
depending on the host system. This configuration exploits vendor- 
or site-supplied MPI installations that often use host-specific, optimized 
network drivers. MPI is thus shown as a virtual dependency in the figure.
ARES builds its own Python version, so that it can run on machines
where Python is not well supported, like Blue~Gene/Q.  Specifically, for
Blue~Gene/Q, it builds Python 2.7.9, which the native software stack
does not support.

Prior to using Spack, ARES managed its software stack with MixDown.
Thus, the ARES team already had some experience supporting 
automated builds of dependencies. We developed Spack packages for the 
packages in Figure~\ref{fig:ares}. Many of the external packages were already 
available in Spack, but some, such as Python, required modifications to support
the new platforms and compilers.

@@

Altogether, the initial packaging effort required roughly two months 
for an experienced build engineer working 20 hours per week. This includes
time spent learning to use Spack.
As shown in the table, 36 different configurations have been run using Spack
(up to 4 versions on each of 10 architecture-compiler-MPI combinations).
Prior to using Spack, only Linux/Intel configurations were automated. The ARES 
team listed a number of key features that enabled the increased automation:
\begin{enumerate}
\item Spack's version tracking and optional dependencies were required to
      build the four configurations with correct libraries;
\item The spec syntax allowed build scripts to test compiler,
      compiler version, and dependency versions concisely---a necessity
      for handling the different architectures;
\item Patching packages for particular platforms was
      necessary to build many packages; and
\item Using a DSL embedded in Python was a significant benefit;
      certain packages required custom scripting to patch.
\end{enumerate}

### 2015-08-10 usecase-policy.tex
Minor tweaks, Final version.

Spack makes it easy to create and to organize package installations,
but it must also be easy for end-users to find and to use packages on an HPC
system. Different end-users have different expectations about how packages
should be built and installed, and at LLNL those expectations are shaped
by years of site policies, personal preferences, and lingering legacy
decisions originally made on a DEC VAX.
Spack provides mechanisms that allow users to customize it for the needs
of particular sites and users.

@@

Spack allows the creation of views, which are a symbolic-link based directory 
layout of packages. Views provide a human-readable directory layout that
can be adapted to match legacy directory layouts on which users rely. For 
example, the above {\tt mpileaks} package may have a view that creates a link 
in {\tt /opt/mpileaks-1.0-openmpi} to the Spack installation of {\tt mpileaks}.
The same package install may be referenced by multiple links and views,
so the above package could also be linked from a more generic {\tt /opt/mpileaks-openmpi}
link. This reuse supports users who always want to use the latest version.
Views can also be used to create symbolic links to specific executables or libraries in an install,
so a Spack-built {\tt gcc@4.9.2} install may have a view that creates links from
{\tt /bin/gcc49} and {\tt /bin/g++49} to the appropriate {\tt gcc} and {\tt g++} executables.

@@

Spack's views are a projection from a point in a high-dimensional space
(the concretized spec, which fully specifies all parameters) to point
in a lower-dimensional space
(the link name, which may only contain a few parameters).
Thus, several installations may map to the same link.
For example, the above {\tt mpileaks} link could point to an {\tt mpileaks} compiled with
{\tt gcc} or {\tt icc} -- the compiler parameter is not part of the link.
To keep  package installations consistent and reproducible,
Spack has a well-defined mechanism for resolving conflicting links;
it uses a combination of internal default policies and user- or site-defined
policies to define an {\it order} of preference for different parameters.
By default, Spack prefers newer versions of packages compiled with newer compilers
to older packages built with older compilers. It has a well-defined, but not
necessarily meaningful, order of preference for deciding between MPI
implementations and different compilers.
The default policies can be overridden in configuration files, by either users
or by sites. For example, at one site users may typically use the Intel compiler,
but some users also use the system's default {\tt gcc@4.4.7}.
These preferences could be stated by adding:
%
\begin{minted}[fontsize=\small]{bash}
    compiler_order = icc,gcc@4.4.7
\end{minted}
%
to the site's configuration file, which would cause the ambiguous
{\tt mpileaks} link to point to an installation compiled with {\tt icc}.
Any compiler not in the {\tt compiler\_order} setting is less 
preferred than those explicitly provided.
%
In a similar manner, Spack can be configured to give specific package
configurations priority over others, which can be useful with a new, unstable 
and untested version.

@@

To support our own private packages, and to support those of LLNL code teams,
Spack allows the creation of site-specific variants of packages.
Users can specify additional search directories for finding additional 
{\tt Package} classes via configuration files.
%
The additional packages are like the {\tt mpileaks} package shown in 
Figure~\ref{fig:mpileaks}.  However, the extension packages can extend from not
only {\tt Package}, but also any of Spack's built-in packages.
Custom packages can inherit from and replace Spack's default
packages, so other sites can either tweak or completely replace Spack's
build recipes. For example, a site can write a local Python class that 
inherits from Spack's base class. The local class may simply add configure 
flags to the Spackclass, while leaving the dependencies and most of the 
build instructions from its parent class. For reproducibility, Spack also 
tracks the {\tt Package} class that drove a specific build.

>>>>

Spack makes it easy to create and to organize package installations,
but it must also be easy for end-users to find and use packages on an HPC
system. Different end-users have different expectations about how packages
should be built and installed, and at LLNL those expectations are shaped
by years of site policies, personal preferences, and lingering legacy
decisions originally made on a DEC VAX.
Spack provides mechanisms that allow users to customize it for the needs
of particular sites and users.

@@

Spack allows the creation of {\it views}, which are symbolic-link based directory 
layouts of packages. Views provide a human-readable directory layout that
can be adapted to match legacy directory layouts on which users rely. For 
example, the above {\tt mpileaks} package may have a view that creates a link 
in {\tt /opt/mpileaks-1.0-openmpi} to the Spack installation of {\tt mpileaks}.
The same package install may be referenced by multiple links and views,
so the above package could also be linked from a more generic {\tt /opt/mpileaks-openmpi}
link. This reuse supports users who always want to use the latest version.
Views can also be used to create symbolic links to specific executables or libraries in an install,
so a Spack-built {\tt gcc@4.9.2} install may have a view that creates links from
{\tt /bin/gcc49} and {\tt /bin/g++49} to the appropriate {\tt gcc} and {\tt g++} executables.

@@

Spack's views are a projection from points in a high-dimensional space
(concretized specs, which fully specify all parameters) to points
in a lower-dimensional space
(link names, which may only contain a few parameters).
Several installations may map to the same link.
For example, the above {\tt mpileaks} link could point to an {\tt mpileaks} compiled with
{\tt gcc} or {\tt icc}---the compiler parameter is not part of the link.
To keep  package installations consistent and reproducible,
Spack has a well-defined mechanism for resolving conflicting links;
it uses a combination of internal default policies and user- or site-defined
policies to define an {\it order} of preference for different parameters.
By default, Spack prefers newer versions of packages compiled with newer compilers
to older packages built with older compilers. It has a well-defined, but not
necessarily meaningful, order of preference for deciding between MPI
implementations and different compilers.
The default policies can be overridden in configuration files, by either users
or by sites. For example, at one site users may typically use the Intel compiler,
but some users also use the system's default {\tt gcc@4.4.7}.
These preferences could be stated by adding:
%
\begin{minted}[fontsize=\small]{bash}
    compiler_order = icc,gcc@4.4.7
\end{minted}
%
to the site's configuration file, which would cause the ambiguous
{\tt mpileaks} link to point to an installation compiled with {\tt icc}.
Any compiler not in the {\tt compiler\_order} setting is less 
preferred than those explicitly provided.
%
In a similar manner, Spack can be configured to give specific package
configurations priority over others, which can be useful with a new, unstable 
and untested version.

@@

To support our own private packages, and to support those of LLNL code teams,
Spack allows the creation of site-specific variants of packages.
Users can specify additional search directories for finding additional 
{\tt Package} classes via configuration files.
%
The additional packages are like the {\tt mpileaks} package shown in 
Figure~\ref{fig:mpileaks}.  However, the extension packages can extend from not
only {\tt Package}, but also any of Spack's built-in packages.
Custom packages can inherit from and replace Spack's default
packages, so other sites can either tweak or completely replace Spack's
build recipes. For example, a site can write a local Python class that 
inherits from Spack's base class. The local class may simply add configure 
flags to Spack's version, while leaving the dependencies and most of the 
build instructions from its parent class. For reproducibility, Spack also 
tracks the {\tt Package} class that drove a specific build.

### 2015-08-10 usecase-python.tex
Minor tweaks, Final version.

\subsection{Support for interpreted languages}
\label{sec:usecase-python}

@@

Per Spack's design philosophy, we wanted a way to manage many different
versions easily, but {\it also} to provide a baseline set of extensions 
{\it without} requiring environment settings.
%
To support this mode of operation, we added the concept of {\tt extension} packages
to Spack. Python modules use the {\tt extends('python')} directive instead of
{\tt depends\_on('python')}.
Each module installs into its own prefix like any other package,
and each module depends on a particular Python installation.
However extensions can be {\tt activated} or {\tt deactivated}
within the dependent Python installation.  The {\tt activate} operation
symbolically links each file in the extension prefix into the Python
installation prefix, as if it were installed directly. If any file
conflict would arise from this operation, {\tt activate} fails.
Similarly, the {\tt deactivate} operation removes the symbolic links and restores
the Python installation to its  pristine state.

@@

Spack essentially implements a ``meta package-manager'' for each Python
instance, which can coexist with Spack's normal installation model.
This mechanism efficiently supports our application teams, for whom we can
now rapidly construct custom Python installations.  We have also reduced
the time that LC staff spend installing Python modules.
Because Spack packages can extend the activation and deactivation mechanisms,
the same mechanism could be used with other interpreted languages with 
similar extension models, such as R, Ruby, or Lua.

>>>>

\subsection{Support for Interpreted Languages}
\label{sec:usecase-python}

@@

Per Spack's design philosophy, we wanted a way to manage many different
versions easily, but {\it also} to provide a baseline set of extensions 
{\it without} requiring environment settings.
%
To support this mode of operation, we added the concept of {\tt extension} packages
to Spack. Python modules use the \verb|extends('python')| directive instead of
\verb|depends_on('python')|.
Each module installs into its own prefix like any other package,
and each module depends on a particular Python installation.
However extensions can be {\tt activated} or {\tt deactivated}
within the dependent Python installation.  The {\tt activate} operation
symbolically links each file in the extension prefix into the Python
installation prefix, as if it were installed directly. If any file
conflict would arise from this operation, {\tt activate} fails.
Similarly, the {\tt deactivate} operation removes the symbolic links and restores
the Python installation to its  pristine state.

@@

Spack essentially implements a ``meta package-manager'' for each Python
instance, which coexists with Spack's normal installation model.
This mechanism efficiently supports application teams, for whom we can
now rapidly construct custom Python installations.  We have also reduced
the time that LC staff spend installing Python modules.
Because Spack packages can extend activation and deactivation mechanisms,
this design could also be used with other languages
with similar extension models, such as R, Ruby, or Lua.
